R TUTORIAL (IV)
========================================================
```{r echo=FALSE, set-options }
options(width = 320)
```
I have not used R for the creation of the network form the bottom-up, but we may use it for plotting and exploration (as Python) and more importantly, for model building and model testing.  
We first need to retrieve the two files we created in Python:
```{r}
library(RCurl)
WarsForRGit <- getURL("https://raw.github.com/MAGALLANESJoseManuel/SocialScienceDataTools/master/TemplatesPython/WarsForR.csv")
el <- read.csv(text = WarsForRGit,header=T)  #list of edges

WarsForR_attrGit <- getURL("https://raw.github.com/MAGALLANESJoseManuel/SocialScienceDataTools/master/TemplatesPython/WarsForR_attr.csv
")
at <- read.csv(text = WarsForR_attrGit,header=T,as.is=T)  #node attributes
```

Let's take a quick look at igraph. 
```{r}
library (igraph)
```

Now let's create a network with igraph:
```{r fig.height=8,fig.width=15}
# el<-read.csv("WarsForR.csv",header=T) #list of edges
# at<-read.csv("WarsForR_attr.csv", header=T, as.is=T) #node attributes
warsGraphIgraph2=graph.data.frame(el,directed=FALSE,vertices=at)
x=V(warsGraphIgraph2)$longitude
y=V(warsGraphIgraph2)$latitude
l2=cbind(x,y)
plot.igraph(warsGraphIgraph2,layout=l2,edge.width=E(warsGraphIgraph2)$duration) #width of edge based on duration!
```
Once we have the network in igraph we can continue using all its tools. Particularly, I recommend reviewing the documentation found at [igraph wiki](http://igraph.sourceforge.net/igraphbook/.

However, R will become extremely important to do network modelling, particularly **ERGM** or **TERGM** models using the **statnet**. 

```{r message=FALSE}
library(statnet)
```

The documentation of statnet is available in  its [website](http://statnet.org/). However, here we will give you the first and most important step!... read a network:
```{r}
# el<-read.csv("WarsForR.csv",header=T)
# at<-read.csv("WarsForR_attr.csv", header=T, as.is=T) #node attributes
```

As we know, the attributes of the nodes consists of only its coordinates. But to vuild a model it is better that we have other kind of attributes, so let's get back our terrorism score, and merge it with the attribute data:
```{r}
path="~/Documents/GITHUBrepositories/Tutorials/TemplatesR/dataCountries.csv"
RegionCode=read.csv(path)[,c(1,3)]
names(RegionCode)
RegionCode$Region=as.numeric(RegionCode$Region)
str(RegionCode$Region)
at=merge(at,RegionCode,by.x="country",by.y="Country",all.x=T)
summary(at)
```

For sure there will be some some countries without this score (i.e. Soviet Union), but let's keep it simple now. Anyway, before using this merge, let's check what we have and get rid of what we do not need:
```{r}
names(at) #variables after the merge
```

Now we can create our network in statnet:
```{r}
nw<-network(el, #edgelist
            matrix.typ="edgelist",          #format of 'el'
            vertex.attr=at,                 #where the node attributes are
            vertex.attrnames=colnames(at),  #names of node attributes
            ignore.eval=FALSE,
            names.eval=c("fatalities","duration","name"),#names of edge attributes
            directed=F, hyper=F, loops=F, multiple=F, bipartite=F)
```

And can see it our statnet network here (two versions):
```{r}
gplot(nw)
gplot(nw,displaylabels=TRUE,usearrows = F)
lnw=cbind(nw %v% "longitude",nw %v% "latitude")
```
...and of course:
```{r fig.height=15,fig.width=18}
gplot(nw,displaylabels=TRUE,coord=lnw,usearrows = F,edge.lwd=nw %e% "duration"/2)
```

Finally, we can show you how you can model a network in statnet:
```{r}
whyWars<-ergm(nw~edges+nodematch("Region"))
```
This model wants to know, for instance, how the terrorism index increases the probability of a link formation. We can see the result below:
```{r}
summary(whyWars)
```
```{r}
simWhyWars <- simulate(whyWars)
plot(simWhyWars)
```
There are many more steps involved in network modelling, and a particularly helpful tutorial on ergm can also be found [here](http://statnet.csde.washington.edu/NME2013/day2/ergm%20tutorial.html).
